Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-05 submission.tex
endringer i sec:data

\section{The Norwegian Dependency Treebank}
\label{sec:data}
We used the newly developed Norwegian Dependency Treebank (NDT)
\cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
Norwegian. It was developed at the National Library of Norway in
collaboration with the University of Oslo, and contains manual
syntactic and morphological annotation. The treebank contains data
from both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
tokens of Nynorsk. We will in the following only be using the Bokmål
portion of the treebank. The annotated texts are mostly newspaper
text, but also include government reports, parliament transcripts and
excerpts from blogs. The annotation process of the treebank was supported by
the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually
corrected by annotators, adding syntactic dependency analysis to the
morphosyntactic annotation.

\paragraph{Morphological Annotation}
The morphological annotation and PoS tag set of NDT follows the Oslo-Bergen
Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work
of \newcite{Faa:Lie:Van:97}. The tag set consists of 12 morphosyntactic PoS
tags outlined in Table \ref{ndttagset}, with 7 additional tags for punctuation
and symbols. The tag set is thus rather coarse-grained, with broad categories
such as \texttt{subst} (noun) and \texttt{verb} (verb). The PoS tags are
complemented by a large set of morphological features, providing information
about morphological properties such as definiteness, number and tense. These
features are used in our tag set modifications, where the coarse PoS tag of
relevant tokens is concatenated with one or more of these features to include
more linguistic information in the tags.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Tag} & \textbf{Description} \\
        \midrule
        \texttt{adj} & Adjective \\
        \texttt{adv} & Adverb \\
        \texttt{det} & Determiner\\
        \texttt{inf-merke} & Infinitive marker \\
        \texttt{interj} & Interjection \\
        \texttt{konj} & Conjunction \\
        \texttt{prep} & Preposition \\
        \texttt{pron} & Pronoun \\
        \texttt{sbu} & Subordinate conjunction \\
        \texttt{subst} & Noun \\
        \texttt{ukjent} & Unknown (foreign word) \\
        \texttt{verb} & Verb \\
        \bottomrule
    \end{tabular}
    \caption{Overview of the PoS tag set of NDT.}
    \label{ndttagset}
\end{table}

\paragraph{Syntactic Annotation}
The syntactic annotation choices in NDT are largely based on the Norwegian
Reference Grammar \cite{Faa:Lie:Van:97}. The annotation choices are outlined in
Table \ref{ndtannotation}, taken from/courtesy of \newcite{Sol:Skj:Ovr:14},
providing overview of the analyses of syntactic constructions that often
distinguish dependency treebanks, such as coordination and the treatment of
auxiliary and main verbs. The set of dependency relations comprises 29
dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
(subject) and \textsc{koord} (coordination).

>>>>

\section{The Norwegian Dependency Treebank}
\label{sec:data}
Our experiments our based on the newly developed Norwegian Dependency Treebank (NDT)
\cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
Norwegian. It was developed at the National Library of Norway in
collaboration with the University of Oslo, and contains manually coded 
syntactic and morphological annotation for 
both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
tokens of Nynorsk. This paper only report results for Bokmål, the main variant.
The treebanked material mostly comprise newspaper
text, but also include government reports, parliament transcripts and
blog excerpts. The annotation process was supported by
the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually
corrected by annotators, adding syntactic dependency analysis to the
morphosyntactic annotation.

\paragraph{Morphological Annotation}
The morphological annotation and PoS tag set of NDT is based on the same
inventory as used by the Oslo-Bergen
Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work
of \newcite{Faa:Lie:Van:97}. The tag set consists of 12 morphosyntactic PoS
tags outlined in Table \ref{ndttagset}, with 7 additional tags for punctuation
and symbols. The tag set is thus rather coarse-grained, with broad categories
such as \texttt{subst} (noun) and \texttt{verb} (verb). The PoS tags are
complemented by a large set of morphological features, providing information
about morphological properties such as definiteness, number and tense. 
Selected subsets of these
features are used in our tag set modifications, where the coarse PoS tag of
relevant tokens is concatenated with one or more features to include 
more linguistic information in the tags.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Tag} & \textbf{Description} \\
        \midrule
        \texttt{adj} & Adjective \\
        \texttt{adv} & Adverb \\
        \texttt{det} & Determiner\\
        \texttt{inf-merke} & Infinitive marker \\
        \texttt{interj} & Interjection \\
        \texttt{konj} & Conjunction \\
        \texttt{prep} & Preposition \\
        \texttt{pron} & Pronoun \\
        \texttt{sbu} & Subordinate conjunction \\
        \texttt{subst} & Noun \\
        \texttt{ukjent} & Unknown (foreign word) \\
        \texttt{verb} & Verb \\
        \bottomrule
    \end{tabular}
    \caption{Overview of the original PoS tag set of NDT (excluding punctuation tags).}
    \label{ndttagset}
\end{table}

\paragraph{Syntactic Annotation}
The syntactic annotation choices in NDT are largely based on the Norwegian
Reference Grammar \cite{Faa:Lie:Van:97}. Some central annotation choices are outlined in
Table \ref{ndtannotation}, taken from of \newcite{Sol:Skj:Ovr:14},
providing overview of the analyses of syntactic constructions that often
distinguish dependency treebanks, such as coordination and the treatment of
auxiliary and main verbs. The annotations comprise 29
dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
(subject) and \textsc{koord} (coordination).
